---
title: "Homework 2"
author: Christian Kroiss
date: 2020-03-30
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(bnlearn)
library(Rgraphviz)
```



```{r}
dag <- model2network("[A][S][E|A:S][O|E][R|E][T|O:R]")
```

```{r}
graphviz.plot(dag)
```

```{r}
d_sep <- bnlearn:::dseparation

d_sep(bn = dag, x = 'O', y = 'R', z=c('E', 'T'))
```

```{r}
vars <- nodes(dag)
pairs <- combn(x = vars, 2, list)
arg_sets <- list()

for (pair in pairs) {
  others <- setdiff(vars, pair)
  conditioning_sets <- unlist(lapply(0:4, function(.x) combn(others, .x, list)), recursive = F)
  for (set in conditioning_sets) {
    args <-list(x = pair[1], y = pair[2], z = set)
    arg_sets <- c(arg_sets, list(args))
  }
}
```

## 1.a

```{r}
d_separated = list()
for (args in arg_sets) {
  if (d_sep(bn=dag, x=args$x, y=args$y, z=args$z)) {
    d_separated <- c(d_separated, list(args))
    print(sprintf("%s and %s are d-separated given {%s}", args$x, args$y, paste(args$z, collapse=",")))
  }
}
```

## 1.b

```{r}
"%subsetof%" <- function(u,v) {
  return(setequal(intersect(u,v),u))
}

entails <- function(rule1, rule2) {
  return((rule1$x == rule2$x) && (rule1$y == rule2$y)
         && (rule1$z %subsetof% rule2$z))
}
str_dsep <- function(ds) {
  sprintf("%s _|_ %s | {%s}", ds$x, ds$y, paste(ds$z, collapse=","))
}
```

```{r}
d_separated_non_redundant <- list()

for (ds in d_separated) {
  print(str_dsep(ds))
  other_ds <- setdiff(d_separated, list(ds))
  entailing <- sapply(other_ds, function(ds2) {entails(ds2, ds)})
  if (!any(entailing)) {
    d_separated_non_redundant <- c(d_separated_non_redundant,
                                   list(ds))
    print(sprintf("No rule entails %s", str_dsep(ds)))
  } else {
    for (ds2 in other_ds[entailing]) {
     print(sprintf("%s entails %s", str_dsep(ds2), str_dsep(ds)))
    }
  }
  print("-------")
}
```

## 1.c

- check subsets in first loop?

## 1.d
```{r}
survey = read.csv("survey2.txt", header = TRUE, sep = " ")
#test_outcome <- ci.test('T', 'E', c('O', 'R'), survey)
test_outcome <- ci.test('A', 'R', z=c('E'), data=survey)
print(test_outcome)
```
```{r}
for (ds in d_separated) {
 
  res <- if (length(ds$z) > 0) {
    ci.test(x = ds$x, y=ds$y, z=ds$z, data=survey)
  } else {
    ci.test(x = ds$x, y=ds$y, data=survey)
  }
  if (res$p.value <= 0.05) {
    print(str_dsep(ds))
    print(res$p.value)
    print("-----")
  }
}
```

## 1.e

All d_separated statements where the independece hypothesis is rejected are redundant.


## 2.a

```{r}
cond_ind = list()
for (args in arg_sets) {
  res <- if (length(args$z) > 0) {
    ci.test(x = args$x, y=args$y, z=args$z, data=survey)
  } else {
    ci.test(x = args$x, y=args$y, data=survey)
  }
  if (res$p.value >= 0.05) {
    cond_ind <- c(cond_ind, list(args))
  }
  
}

```

## 2.b
What is the proportion of true D-separation
statements that are also true conditional independence statements?

```{r}
p_d_sep_ci <- (length(intersect(cond_ind, d_separated)) / 
  length(d_separated))
print(p_d_sep_ci)

```

## 2.c
What is the proportion of true conditional independence statements that are also true-deseparation statements?

```{r}
p_ci_dsep <- (length(intersect(cond_ind, d_separated)) / 
  length(cond_ind))
print(p_ci_dsep)
```

## 2.d
How would these results change if we only considered non-redundant d-separation statements?

```{r}
(length(intersect(cond_ind, d_separated_non_redundant)) / 
  length(d_separated_non_redundant))
```
```{r}
(length(intersect(cond_ind, d_separated_non_redundant)) / 
  length(cond_ind)) 
```
## 2.e
**Q:** Based on these results, how well do the faithfulness assumption and global Markov property assumption hold
up with this DAG and dataset?


**A:** global Markov property holds for non-redundant d-separation statements. Faithfulness doesn't hold

