---
title: "Homework 2"
author: Christian Kroiss
date: 2020-03-30
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(bnlearn)
library(Rgraphviz)
```



```{r}
dag <- model2network("[A][S][E|A:S][O|E][R|E][T|O:R]")
```

```{r}
graphviz.plot(dag)
```

```{r}
d_sep <- bnlearn:::dseparation

d_sep(bn = dag, x = 'O', y = 'R', z=c('E', 'T'))
```

```{r}
vars <- nodes(dag)
pairs <- combn(x = vars, 2, list)
arg_sets <- list()

for (pair in pairs) {
  others <- setdiff(vars, pair)
  conditioning_sets <- unlist(lapply(0:4, function(.x) combn(others, .x, list)), recursive = F)
  for (set in conditioning_sets) {
    args <-list(x = pair[1], y = pair[2], z = set)
    arg_sets <- c(arg_sets, list(args))
  }
}
```

## 1.a

```{r}
d_separated = list()
for (args in arg_sets) {
  if (d_sep(bn=dag, x=args$x, y=args$y, z=args$z)) {
    d_separated <- c(d_separated, list(args))
    print(sprintf("%s and %s are d-separated given {%s}", args$x, args$y, paste(args$z, collapse=",")))
  }
}
```

## 1.b

```{r}
"%subsetof%" <- function(u,v) {
  return(setequal(intersect(u,v),u))
}

entails <- function(rule1, rule2) {
  return((rule1$x == rule2$x) && (rule1$y == rule2$y)
         && (rule1$z %subsetof% rule2$z))
}
str_dsep <- function(ds) {
  sprintf("%s _|_ %s | {%s}", ds$x, ds$y, paste(ds$z, collapse=","))
}
```

```{r}
d_separated_non_redundant <- list()

for (ds in d_separated) {
  print(str_dsep(ds))
  other_ds <- setdiff(d_separated, list(ds))
  entailing <- sapply(other_ds, function(ds2) {entails(ds2, ds)})
  if (!any(entailing)) {
    d_separated_non_redundant <- c(d_separated_non_redundant,
                                   list(ds))
    print(sprintf("No rule entails %s", str_dsep(ds)))
  } else {
    for (ds2 in other_ds[entailing]) {
     print(sprintf("%s entails %s", str_dsep(ds2), str_dsep(ds)))
    }
  }
  print("-------")
}
```

## 1.c

- check subsets in first loop?

## 1.d
```{r}
survey = read.csv("survey2.txt", header = TRUE, sep = " ")
#test_outcome <- ci.test('T', 'E', c('O', 'R'), survey)
test_outcome <- ci.test('A', 'R', z=c('E'), data=survey)
print(test_outcome)
```
```{r}
for (ds in d_separated) {
 
  res <- if (length(ds$z) > 0) {
    ci.test(x = ds$x, y=ds$y, z=ds$z, data=survey)
  } else {
    ci.test(x = ds$x, y=ds$y, data=survey)
  }
  if (res$p.value <= 0.05) {
    print(str_dsep(ds))
    print(res$p.value)
    print("-----")
  }
}
```

## 1.e

All d_separated statements where the independece hypothesis is rejected are redundant.


## 2.a

```{r}
cond_ind = list()
for (args in arg_sets) {
  res <- if (length(args$z) > 0) {
    ci.test(x = args$x, y=args$y, z=args$z, data=survey)
  } else {
    ci.test(x = args$x, y=args$y, data=survey)
  }
  if (res$p.value >= 0.05) {
    cond_ind <- c(cond_ind, list(args))
  }
  
}

```

## 2.b
What is the proportion of true D-separation
statements that are also true conditional independence statements?

```{r}
p_d_sep_ci <- (length(intersect(cond_ind, d_separated)) / 
  length(d_separated))
print(p_d_sep_ci)

```

## 2.c
What is the proportion of true conditional independence statements that are also true-deseparation statements?

```{r}
p_ci_dsep <- (length(intersect(cond_ind, d_separated)) / 
  length(cond_ind))
print(p_ci_dsep)
```

## 2.d
How would these results change if we only considered non-redundant d-separation statements?

```{r}
(length(intersect(cond_ind, d_separated_non_redundant)) / 
  length(d_separated_non_redundant))
```
```{r}
(length(intersect(cond_ind, d_separated_non_redundant)) / 
  length(cond_ind)) 
```
## 2.e
**Q:** Based on these results, how well do the faithfulness assumption and global Markov property assumption hold
up with this DAG and dataset?


**A:** global Markov property holds for non-redundant d-separation statements. Faithfulness doesn't hold.

## 3.
```{r}
net <- model2network('[A][B|A][C|B:A]')
nombres <- c('off', 'on')
cptA <- matrix(c(0.5, 0.5), ncol=2)
dimnames(cptA) <- list(NULL, nombres)
cptB <- matrix(c(.8, .2, .1, .9), ncol=2)
dimnames(cptB) <- list(B = nombres, A = nombres)
cptC <- matrix(c(.9, .1, .99, .01, .1, .9, .4, .6))
dim(cptC) <- c(2, 2, 2)
dimnames(cptC) <- list(C = nombres, A = nombres, B = nombres)
model <- custom.fit(net, list(A = cptA, B = cptB, C = cptC))
graphviz.plot(model)
```
## 3.a

```{r}
#model$C
dim(model$C$prob)
p_res = 0.5 * 0.9 * 0.6 / (0.5 * 0.9 * 0.6 + 0.5 * 0.2 * 0.9)
print(p_res)
```
$P(A=on|B=on, C=on) = \frac{P(A=on)P(B=on|A=on)P(C=on|A=on,B=on)}{P(A=on)P(B=on|A=on)P(C=on|A=on,B=on) + P(A=off)P(B=on|A=off)P(C=on|A=off,B=on)}$

$= \frac{0.5 * 0.9 * 0.6}{0.5 * 0.9 * 0.6 + 0.5 * 0.2 * 0.9}$

$= 0.75$

## 3.b
```{r}
sampled <- rbn(model, n=100000)
```

```{r}
sampled_b1_c1 <- sampled %>% filter((B == "on") & (C == "on"))

sampled_a1_b1_c1 <- sampled_b1_c1 %>% filter(A == "on")

sampled_a0_b1_c1 <- sampled_b1_c1 %>% filter(A == "off")
```

```{r}
p_rejection_sampling <- dim(sampled_a1_b1_c1)[1] / dim(sampled_b1_c1)[1]
print(p_rejection_sampling)
```

## 3.c

```{r}
model_mutilated <- mutilated(model, list(B="on"))
graphviz.plot(model_mutilated)

```
```{r}
model_mutilated$B
```

## 3.d
```{r}
p_res_mutilated = 0.5 * 1 * 0.6 / (0.5 * 1 * 0.6 + 0.5 * 1 * 0.9)
print(p_res_mutilated)
```

## 3.e
```{r}
sampled_mutilated <- rbn(model_mutilated, n=100000)
```

```{r}
sampled_gm_b1_c1 <- sampled_mutilated %>% filter((B == "on") & (C == "on"))

sampled_gm_a1_b1_c1 <- sampled_gm_b1_c1 %>% filter(A == "on")

```

```{r}
p_gm_rejection_sampling <- dim(sampled_gm_a1_b1_c1)[1] / dim(sampled_gm_b1_c1)[1]
print(p_gm_rejection_sampling)
```

```

