---
title: "Homework 1"
author: Christian Kroiss
date: 2020-03-18
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(bnlearn)
library(Rgraphviz)
```

## 1. Building a DAG

$P(A, S, E, O, R, T) = P(A)P(S)P(E|A,S)P(O|E)P(R|E)P(T|O,R)$


```{r}
dag <- model2network("[A][S][E|A:S][O|E][R|E][T|O:R]")
```

Class: `r class(dag)`


```{r}
graphviz.plot(dag)
```

## 2. Experimenting with graph utilities

### Nodes
```{r}
dag$nodes$A
```
### Arcs
```{r}
dag$arcs
```

### a,b,c) Parents, Children, Markov Blanket
```{r}
for (n in names(dag$nodes)) {
  print(sprintf("%s: parents = %s, children = %s, markov blanket = %s", n, 
                paste(parents(dag, n), collapse = ", "),
                paste(children(dag, n), collapse = ", "),
                paste(mb(dag, n), collapse = ", ")))
}
```

### d) Markov blanket

Let $X = \{A, E, O, R, S, T \}$ be the set of variables in the joint distribution $P(A, E, O, R, S, T)$. Then, a markov blanket of a variable $x \in X$ is a subset $B \subset X$ for which the following holds: 
given $B$, then $x$ is conditionally independent of all other variables. i.e.  
$$
P(x| X \setminus \{x\}) = P(x| B)
$$

### e) Identifying markov blankets

For finding a markov blanket for variable $x$, I check every path to every other variable in the graph and determine which nodes have to be blocked to block the path in terms of d-separation.

## 3. CPD estimation

```{r}
survey = read.csv("survey2.txt", header = TRUE, sep = " ")
model <- bn.fit(dag, data = survey, method = "bayes")
```
```{r}
model
```



## 4, Graph manipulation
```{r}
dag2 <- dag
dag2 <- drop.arc(dag2, "E", "O")
graphviz.plot(dag2)
```

```{r}
model2 <- bn.fit(dag2, data = survey, method = "bayes")
```

```{r}
model$O$prob
```

```{r}
cpquery(model, (O == "emp"), (E == "high"))
```


```{r}
model2$O$prob
```

Only the CPT of `O` changed. In the model after the graph manipulation, the CPT for `O` has only 2 entries instead of 4. The probabilities for `emp` and `self` are close to the averages of the respective rows from the original model, containing entries for `high` and `uni` of `E`.

## 5. Markov equivalence
```{r}
P1 = cpdag(dag)
```

```{r fig.cap="P1"}
graphviz.plot(P1)
```


```{r}
D2 <- dag
D2 <- set.arc(D2, "O", "R")
```

```{r}
graphviz.plot(D2)
```
```{r}
P2 <- cpdag(D2)
```

```{r}
graphviz.plot(P2)
```

```{r}
P1$arcs
```
